Improved Bayesian Ethnorace Prediction

### 2021-04-28 paper.Rmd
made plots grayscale :(

```{=tex}
\begin{singlespace}
Quantitative social science research on race and ethnicity can be constrained by the lack of available individual-level data with these markers. In response to this issue, methods have been developed to predict individuals' race and ethnicity using Bayes' rule. I expand upon existing methods by incorporating information into the algorithm. Along with some other adjustments, these improvements lead to substantial gains in predictive performance when validated against official state voter files. The largest of these predictive gains are found for African Americans---a group which existing methods find difficult to predict accurately. Furthmore, I apply my algorithm to previous reseach using the old methods and demonstrate the substantive implications of switching to more accurate predictions.
\end{singlespace}
```
# Introduction

@@

In the next section I provide some background and specifics regarding the inputs and outputs of my ethnorace prediction method. Then I explain the methodology and compare my implementation with previous versions. Finally, I demonstrate the predictive performance of my method when validated against the combined North Carolina and Florida voter file ($n$ = 21,164,503).

@@

There are a few benefits to using the Census ethnorace categorization. These definitions capture a common understanding of race and ethnicity in the US, and correspond to the groups found most frequently in social science research. The data sources of these groups' distributions that serve as inputs to the prediction formula also rely on the Census categorization. This also facilitates comparison of my method against previous ethnorace prediction methods.[^3] One downside to using the Census categories, however, is that it obscures substantial heterogeneity that may exist within each group. Within Asian Americans and Latinos, for example, there is considerable variation in terms of national ancestry. Furthermore, the unfortunate necessity of a catch-all Other Race category ensures that important sources of diversity are lost.[^4]

@@

The first names list I use comes from Tzioumis (2017). It is drawn from mortgage applications and contains ethnorace counts in each of the six groups across 4,250 first names. Unlike Census data, which form the basis for much of my other data sources, this list of first names may be unrepresentative of the larger US population. To the extent that first name distributions differ by ethnorace given employment status, for example, this may be a concern. But the predictive benefits from using first name data, as I will demonstrate, likely overwhelm these worries in most contexts.

@@

For my last names data, I use the 2010 Census Surnames List.[^5] This list comes from the 2010 decennial Census and contains over 160,000 common US last names (those occurring 100 or more times in the population). Like the first names list, these data include counts of individuals in each of the six ethnorace categories across each last name.

@@

@@

My ethnorace distributions by geographies come from the 2010 decennial Census, accessed via IPUMS NHGIS.[^6] In decreasing order of mean population, these geographies include *state*, *county*, *Census place*, *ZIP code*, and *Census block*. Predictions tend to improve with more precise levels of geography. With this in mind, my implementation automatically matches each individual to the most fine-grain level of geography available.

[^6]: Steven Manson, Jonathan Schroeder, David Van Riper, Tracy Kugler, and Steven Ruggles. IPUMS National Historical Geographic Information System: Version 15.0 [dataset]. Minneapolis, MN: IPUMS. 2020. <http://doi.org/10.18128/D050.V15.0>

@@

My party identification data come from a 2012 Gallup poll.[^7] The three categories of political party I include are Republican, Democrat, and Other (including Independents and "don't knows"). The Gallup report tells me both the probability that an individual with a given ethnorace belongs to a particular political party, and the probability that an individual with a given political party identifies with a particular ethnorace.

[^7]: <https://news.gallup.com/poll/160373/democrats-racially-diverse-republicans-mostly-white.aspx>

@@

These data refer to ethnorace distributions over multi-unit housing occupancy. Individuals are matched to these probabilities if their address contains "Apt", "Unit", "\#", or other such identifier. Unfortunately, I was not able to find these distributions for the 2010 decennial Census, so instead I use those from the year 2000.

@@

The raw data sources I describe above, with the exception of party ID,[^8] all contain counts of individuals with a particular attribute (i.e. the first name JOHN, or the ZIP Code "92092") per ethnorace category. Taking proportions by cell across a given attribute tells us $Pr(Ethnorace|Attribute)$, and taking proportions by cell across a given ethnorace group tells us $Pr(Attribute|Ethnorace)$. These two conditional probabilities form the building blocks of the classification algorithm described later.

[^8]: Party ID percentages by ethnorace are directly available in the Gallup report.

If any cell in the input data is empty (i.e. if there are no individuals of a particular ethnorace with some attribute), then the conditional probabilities $Pr(Ethnorace|Attribute)$ and $Pr(Attribute|Ethnorace)$ will be zero. As will become clear in the Methodology section, if either of those two probabilities for an individual equal zero for a given ethnorace, the algorithm will predict a zero percent probability that the individual belongs to that ethnorace. This will occur even if some other attributes about that individual predict a high probability of belonging to the ethnorace. For example, an individual could have first and last names that are highly predictive of being Hispanic, but reside in a Census block which had zero Hispanic occupants at the time of the 2010 decennial Census. Blocks typically contain only around 400 individuals---so this is a real possibility. For this hypothetical person the input data claims that $Pr(Hispanic|Census Block) = 0$, which yields $Pr(Hispanic)= 0$ due to the structure of the prediction algorithm. To resolve this issue, I apply a technique from the machine learning literature known as Laplace smoothing to my input data. This works by adding some constant, or psuedocount, to number of individuals in every cell in the input data, then calculating the conditional probabilities $Pr(Ethnorace|Attribute)$ and $Pr(Attribute|Ethnorace)$.[^9] Through validation tests, I found that Laplace smoothing led to significant gains in predictive performance for Asian individuals in particular. I also conjecture that, absent this smoothing technique, the algorithm's predictions are too beholden to the specifics of the 2010 decennial Census. The predictions will generalize better to other time periods without the rigid assumptions of zero conditional probability for some attribute/ethnorace combinations.

[^9]: Missing counts are only a problem in the first names, last names, and geolocation data so psuedocounts are only added for those inputs. The exact value for the Laplace smoothing psuedocount could be any number greater than zero, but through out-of-sample validation I found 5 to be the optimal value.

@@

Where $x$ is the vector of individual attributes indexed by $j$. The particular attribute $x'$ comes from using the chain rule to decompose the joint probability $Pr(R=r, X)$. The choice of which attribute to use for $x'$ is atheoretical, but all previous prediction methods have used last names [@elliott2009; @imai2016; @voicu2018]. During my validation exercises, I found that the choice of $x'$ has potentially large consequences for predictive performance. For example, using last names for $x'$ appears to help predictions of Whites---but to the detriment of non-Whites. In light of these trade-offs, my method cycles through every attribute as the choice of $x'$ and computes $Pr(R=r|X)$ for each. These posterior probabilities are then averaged within each ethnoracial category to generate final predicted probabilities that an individual belongs to a particular ethnorace. The result is more balanced predictions across each ethnorace.

The conditional independence assumption necessary for transforming equation (1) to (2) says that knowing both a particular attribute of an individual, and that individual's ethnorace, should give us no extra knowledge of any other attribute for that individual. Stated formally, $Pr(x_j|R=r,X) = Pr(x_j|R=r)$ for all $x_j$. This assumption is almost certainly violated in the present context. One example that has been demonstrated empirically is that last name distributions by race vary across regions in the US [@crabtree2018].

Violations of the conditional independence assumption are commonplace in most applications of similar classification algorithms. Nevertheless, these prediction methods perform unreasonably well in many contexts [@lewis1998; @domingos1997; @rish2001]. This is likely because of the decision rule governing the final classifications---the posterior probabilities of the true class do not have to necessarily be statistically valid, they only need to be higher than those of every other class to be accurately classified. As is true with my method, when more attribute inputs are added to the model (first names, multi-unit occupancy) researchers should be cautious trying to interpret the posterior probabilities $Pr(R=r|X)$ directly. Rather, the maximum a posteriori ethnorace classifications should be used alone.

@@

To test the performance of the model, I apply the predictions to the combined North Carolina and Florida State voter file. These files contain snapshots of the registered voters in their respective states and provide individual-level data for first names, last names, address, political party, age, gender, and crucially self-identified ethnorace. After combining the two voter files, I then geocoded each unique address in the sample. This allowed me to match individual observations to Census places and blocks, and ZIP codes. Then I applied the prediction algorithm described above using the `bper` package and calculated each individuals' predicted ethnorace. In order to compare my method against an existing benchmark, I also used the `wru` package [@imai2016] to calculate ethnorace predictions for the same individuals.

@@

```{r, eval=F, fig.cap="Comparison to True Population Proportions in North Carolina and Florida"}

# 
# true_props <-
#   tibble(
#     prop_api = 0.017,
#     prop_black = 0.163,
#     prop_hispanic = 0.111,
#     prop_other = 0.058,
#     prop_white = 0.651
#   ) %>% 
#   gather(value = true_props)
# 
# wru_props <- tests[1, c("prop_api", "prop_black", "prop_hispanic", "prop_other", "prop_white")] %>% 
#   gather(value = "props") %>% 
#   mutate(Method = "wru")
# 
# props <- tests[2, c("prop_api", "prop_black", "prop_hispanic", "prop_other", "prop_white")] %>% 
#   gather(value = "props") %>%
#   mutate(Method = "bper") %>% 
#   rbind(wru_props) %>% 
#   left_join(true_props) %>% 
#   mutate(diff = props - true_props,
#          key = case_when(key == "prop_api" ~ "Asian",
#                          key == "prop_black" ~ "Black",
#                          key == "prop_hispanic" ~ "Hispanic",
#                          key == "prop_other" ~ "Other",
#                          key == "prop_white" ~ "White"))
# 
# 
# ggplot(props, aes(x = key, y = diff * 100, fill = Method)) +
#   geom_bar(stat = "identity", position = position_dodge()) +
#   theme_minimal() +
#   labs(y = "% Point Difference from True Proportion", x = "") +
#   theme(text = element_text(family = "serif")) +
#   scale_fill_manual(values = c("darkcyan", "brown"))

```

The most straightforward metric for assessing individual-level predictive performance is the Accuracy score, or Overall Error Rate. This number is the proportion of correctly classified individuals in the sample. I ran models separately for different combinations of input variables to mimic data availability constraints in real-world applications, and then calculated the Accuracy score for each. Figure 1 displays a summary of the results of these different models.

@@

ggplot(acc_bper, aes(x = test_type, y = accuracy, group = 1)) +
  geom_line(color = "darkcyan") +
  geom_point(size = 3, color = "darkcyan") +
  theme_minimal() +
  ylim(0.75, 0.85) +
  labs(y = "Accuracy Score") +
  scale_x_discrete(name = "Model Inputs", limits = rev(c("all", "all_np", "zip", "zip_np", "place", "place_np", "county", "county_np", "state", "state_np")),
                   labels = rev(c(
                     "all" = "Block \n Multi-Unit \n Party ID \n First Name \n Last Name", 
                     "place" = "Place \n Party ID \n First Name \n Last Name", 
                     "county" = "County \n Party ID \n First Name \n Last Name", 
                     "all_np" = "Block \n Multi-Unit \n First Name \n Last Name", 
                     "place_np" = "Place \n First Name \n Last Name", 
                     "county_np" = "County \n First Name \n Last Name",
                     "zip" = "ZIP \n Party ID \n First Name \n Last Name",
                     "zip_np" = "ZIP \n First Name \n Last Name",
                     "state" = "State \n Party ID \n First Name \n Last Name",
                     "state_np" = "State \n First Name \n Last Name"
                   ))) +
  theme(text = element_text(family = "serif"), axis.text.x = element_text(size = 7))
```

As expected, the model with the greatest number of input data sources, and at the most precise geographic level, on the far right of the figure performs the best in terms of overall Accuracy. Using Census blocks, multi-unit address, party ID, first names, and last names, this model correctly identifies the ethnorace of `r round(subset(tests, select = "accuracy", test_method == "bper" & test_type == "all"), 3) * 100`% of individuals in the sample. The upward trend in model accuracy seen in the figure corresponds to shrinking the size of the geolocation variable used. Moving from state to county, from county to place, from place to ZIP code, and from ZIP code to block all improve the overall predictive performance in terms of overall Accuracy. I include the input for multi-unit occupancy only for the Census block models because I believe this reflects the practical contexts where `bper` might be used. If a researcher has access to individual-level addresses, they should be able to geocode these to find the matching Census blocks and should also be able to parse the residency type (multi-unit or stand-alone). But researchers relying on more aggregate geographies likely do not have access to individual addresses, and hence the residency characteristics, for their sample.[^10] In Figure 1 I pair each geolocation variable with a model missing the party ID input. The gains to predictive performance across all geographies appears roughly uniform with the addition of party ID.

[^10]: In the event that data is available, adding multi-unit occupancy inputs to aggregate geographies, such as ZIP codes, places, counties, or states *greatly* enhances the predictive accuracy of the model.

@@

ggplot(metrics_bper, aes(x = test_type, y = value, group = Metric)) +
  geom_line(aes(color = Metric), size = 0.3) +
  geom_point(aes(color = Metric), size = 2) +
  theme_minimal() +
  labs(x = "", y = "Score") +
  scale_color_manual(values = c("gold3", "skyblue4")) +
  scale_x_discrete(name = "Model Inputs", limits = rev(c("all", "all_np", "zip", "zip_np", "place", "place_np", "county", "county_np", "state", "state_np")),
                   labels = rev(c(
                     "all" = "Block", 
                     "place" = "Place", 
                     "county" = "County ", 
                     "all_np" = "Block (np)", 
                     "place_np" = "Place (np)", 
                     "county_np" = "County (np)",
                     "zip" = "ZIP",
                     "zip_np" = "ZIP (np)",
                     "state" = "State",
                     "state_np" = "State (np)"
                   ))) +
  facet_wrap(~ Race) +
  theme(text = element_text(family = "serif"), axis.text.x = element_text(size = 7, angle = 60))
```

Figure 2 displays the Precision and Recall scores broken down by ethnorace for the same models used in Figure 1.[^11] For both Whites and Hispanics, Precision and Recall remain high across all models. Using the most input data available (the models on the far right of the figure), White Precision and Recall are `r round(subset(tests, select = "prec_white", test_method == "bper" & test_type == "all"), 3)` and `r round(subset(tests, select = "rec_white", test_method == "bper" & test_type == "all"), 3)`, respectively. Hispanic Precision and Recall in this model are `r round(subset(tests, select = "prec_hispanic", test_method == "bper" & test_type == "all"), 3)` and `r round(subset(tests, select = "rec_hispanic", test_method == "bper" & test_type == "all"), 3)`, respectively. Hispanic predictive performance across all models likely remains high due to the distinctiveness of Spanish surnames. These metrics are uniformly lower for Asians---likely reflecting the relative rarity of Asian individuals in the sample. When using Census blocks as the unit of geography, predictions for African Americans are quite strong (Precision: `r round(subset(tests, select = "prec_black", test_method == "bper" & test_type == "all"), 3)`, Recall: `r round(subset(tests, select = "rec_black", test_method == "bper" & test_type == "all"), 3)`). But predictive performance falls with broader geolocations. At the state level, without party ID, Black Precision is `r round(subset(tests, select = "prec_black", test_method == "bper" & test_type == "state_np"), 3)` and Recall is `r round(subset(tests, select = "rec_black", test_method == "bper" & test_type == "state_np"), 3)`. This difference in predictive performance between geographies could be explained by the legacy of Black segregation in at least the North Carolina and Florida sample.

[^11]: The model names are abbreviated by the level of geography, and all use first name and last name inputs. Models with *(np)* do not use the party ID inputs.

@@

```{r, fig.cap="Probability Calibration Plots by Ethnorace"}
cal_api <-
  ggplot(cal_data[[12]]) +
  geom_line(color = "darkcyan") +
  geom_point(color = "darkcyan") +
  theme_minimal() +
  labs(y = "Observed Asian Percentage",
       x = "Predicted Probability Asian",
       title = "Asian") +
  ylim(0, 100) +
  theme(text = element_text(family = "serif"), plot.title = element_text(hjust = 0.5))

cal_black <-
  ggplot(cal_data[[11]]) +
  geom_line(color = "darkcyan") +
  geom_point(color = "darkcyan") +
  theme_minimal() +
  labs(y = "Observed Black Percentage",
       x = "Predicted Probability Black",
       title = "Black") +
  ylim(0, 100) +
  theme(text = element_text(family = "serif"), plot.title = element_text(hjust = 0.5))

>>>>

```{=tex}
\begin{singlespace}
Quantitative social science research on race and ethnicity can be constrained by the lack of available individual-level data with these markers. In response to this issue, methods have been developed to predict individuals' race and ethnicity using Bayes' rule. I expand upon existing methods by incorporating new information into the algorithm. Along with some other adjustments, these improvements lead to substantial gains in predictive performance when validated against official state voter files. The largest of these predictive gains are found for African Americans---a group which existing methods find difficult to predict accurately. Furthermore, I apply my algorithm to previous research using the old methods and demonstrate the substantive implications of switching to more accurate predictions.
\end{singlespace}
```
# Introduction

@@

In the next section I provide some background and specifics regarding the inputs and outputs of my ethnorace prediction method. Then I explain the methodology and compare my implementation with previous versions. Next, I demonstrate the predictive performance of my method when validated against the combined North Carolina and Florida voter file ($n$ = 21,164,503). And lastly, I replicate Grumbach and Sahn's (2020) descriptive findings in their study on racial disparities in campaign finance using my method of ethnorace prediction. This demonstrates some of the substantive empirical implications of using improved predictions.

@@

There are a few benefits to using the Census ethnorace categorization. These definitions capture a common understanding of race and ethnicity in the US, and correspond to the groups studied most frequently in social science research. The data sources of these groups' distributions that serve as inputs to the prediction formula also rely on the Census categorization. There is currently no feasible alternative data source that I know of that offers the same richness as the Census. This also facilitates comparison of my method against previous ethnorace prediction methods.[^3] One downside to using the Census categories, however, is that it obscures substantial heterogeneity that may exist within each group. Within Asian Americans and Latinos, for example, there is considerable variation in terms of national ancestry. Furthermore, the unfortunate necessity of a catch-all Other Race category ensures that important sources of diversity are lost.[^4] Researchers interested in using this ethnorace prediction method should be aware of these shortcomings. Nonetheless, because the output groups correlate strongly with many important outcomes in social science research, there are still many applications where this prediction method will be useful.

@@

The first names list I use comes from Tzioumis (2017). It is drawn from mortgage applications and contains ethnorace counts in each of the six groups across 4,250 first names. Relative to Census data, this list of first names may be unrepresentative of the larger US population. Mortgage applicants are wealthier than the average American, and are more likely to be employed. To the extent that first name distributions differ by ethnorace given these unobserved characteristics, this may be a concern. But the predictive benefits from using first name data, as I will demonstrate, likely overcome these worries.

@@

For my last names data, I draw from the 2010 Census Surnames List.[^5] This list comes from the 2010 decennial Census and contains over 160,000 common US last names (those occurring 100 or more times in the population). I also append any last names from the 2000 Census Surnames List that are not found in the 2010 list. This increases the probability the algorithm will match an individual's last name. Like the first names list, these data include counts of individuals in each of the six ethnorace categories across each last name. Once the 2020 Census Surnames List is released, these data will be readily incorporated as well.[^6]

@@

[^6]: Because most of the input data sources are from the 2010 decennial Census, the predictions are likely most accurate when applied to data sets from around that time period. The algorithm could be improved in the future by allowing flexibility in which years to draw input data from.

@@

My ethnorace distributions by geographies come from the 2010 decennial Census, accessed via IPUMS NHGIS.[^7] In decreasing order of mean population, these geographies include *state*, *county*, *Census place*, *ZIP code*, and *Census block*. Predictions tend to improve with more precise levels of geography. With this in mind, my implementation automatically matches each individual to the most fine-grain level of geography available. As with the last names input data, once the 2020 Census is released the algorithm will include the updated geolocation distributions.

[^7]: Steven Manson, Jonathan Schroeder, David Van Riper, Tracy Kugler, and Steven Ruggles. IPUMS National Historical Geographic Information System: Version 15.0 [dataset]. Minneapolis, MN: IPUMS. 2020. <http://doi.org/10.18128/D050.V15.0>

@@

My party identification data come from a 2012 Gallup poll.[^8] The three categories of political party I include are Republican, Democrat, and Other (including Independents and "don't knows"). The Gallup report tells me both the probability that an individual with a given ethnorace belongs to a particular political party, and the probability that an individual with a given political party identifies with a particular ethnorace. Party ID by ethnorace has likely fluctuated since 2012, yet the inclusion of these data still modestly improve predictions for recent data sets.

[^8]: <https://news.gallup.com/poll/160373/democrats-racially-diverse-republicans-mostly-white.aspx>

@@

These data refer to ethnorace distributions over multi-unit housing occupancy. Individuals are matched to these probabilities if their address contains "Apt", "Unit", "\#", or other such identifier. Unfortunately, I was not able to find these distributions for the 2010 decennial Census, so instead I use those from the year 2000. Again, hopefully I will be able to incorporate data from the 2020 decennial Census here when, or if, it becomes available.

@@

The raw data sources I describe above, with the exception of party ID,[^9] all contain counts of individuals with a particular attribute (i.e. the first name JOHN, or the ZIP Code "92092") per ethnorace category. Taking proportions by cell across a given attribute tells us $Pr(Ethnorace|Attribute)$, and taking proportions by cell across a given ethnorace group tells us $Pr(Attribute|Ethnorace)$. These two conditional probabilities form the building blocks of the classification algorithm described later.

[^9]: Party ID percentages by ethnorace are directly available in the Gallup report.

If any cell in the input data is empty (i.e. if there are no individuals of a particular ethnorace with some attribute), then the conditional probabilities $Pr(Ethnorace|Attribute)$ and $Pr(Attribute|Ethnorace)$ will be zero. As will become clear in the Methodology section, if either of those two probabilities for an individual equal zero for a given ethnorace, the algorithm will predict a zero percent probability that the individual belongs to that ethnorace. This will occur even if some other attributes about that individual predict a high probability of belonging to the ethnorace. For example, an individual could have first and last names that are highly predictive of being Hispanic, but reside in a Census block which had zero Hispanic occupants at the time of the 2010 decennial Census. Blocks typically contain only around 400 individuals---so this is a real possibility. For this hypothetical person the input data claims that $Pr(Hispanic|Census Block) = 0$, which yields $Pr(Hispanic)= 0$ due to the structure of the prediction algorithm. To resolve this issue, I apply a technique from the machine learning literature known as Laplace smoothing to my input data. This works by adding some constant, or psuedocount, to number of individuals in every cell in the input data, then calculating the conditional probabilities $Pr(Ethnorace|Attribute)$ and $Pr(Attribute|Ethnorace)$.[^10] Through validation tests, I found that Laplace smoothing led to significant gains in predictive performance for Asian individuals in particular. I also conjecture that, absent this smoothing technique, the algorithm's predictions are too beholden to the specifics of the 2010 decennial Census. The predictions will generalize better to other time periods without the rigid assumptions of zero conditional probability for some attribute/ethnorace combinations.

[^10]: Missing counts are only a problem in the first names, last names, and geolocation data so psuedocounts are only added for those inputs. The exact value for the Laplace smoothing psuedocount could be any number greater than zero, but through out-of-sample validation I found 5 to be the optimal value.

@@

Where $x$ is the vector of individual attributes indexed by $j$. The particular attribute $x'$ comes from using the chain rule to decompose the joint probability $Pr(R=r, X)$. The choice of which attribute to use for $x'$ is atheoretical, but all previous prediction methods have used last names [@elliott2009; @imai2016; @voicu2018]. During my validation exercises, I found that the choice of $x'$ has potentially large consequences for predictive performance. This is because the expression $Pr(R=r|x')$ is typically much greater than any $Pr(x_j|R=r)$, and therefore contributes more weight to the final posterior probability.[^11] For example, using last names for $x'$ appears to help predictions of Whites---but to the detriment of non-Whites. In light of these trade-offs, my method cycles through every attribute as the choice of $x'$ and computes $Pr(R=r|X)$ for each. These posterior probabilities are then averaged within each ethnoracial category to generate final predicted probabilities that an individual belongs to a particular ethnorace. The result of this smoothing is more balanced predictions across each ethnorace.

[^11]: To see why, imagine $x = Smith$ and $r = Black$. The probability than an individual is Black given they have the last name Smith, $Pr(R=r|x)$, is 0.23. But the probability that an individual is named Smith given that they are Black, $Pr(x|R=r)$, is 0.0000029.

The conditional independence assumption necessary for transforming equation (1) to (2) says that knowing both a particular attribute of an individual, and that individual's ethnorace, should give us no extra knowledge of any other attribute for that individual. Stated formally, $Pr(x_j|R=r,X) = Pr(x_j|R=r)$ for all $x_j$. This assumption is almost certainly violated in the present context. One example that has been demonstrated empirically is that last name distributions by race vary across regions in the US [@crabtree2018]. Violations of the conditional independence assumption are commonplace in most applications of similar "naive" Bayes classification algorithms. Nevertheless, these prediction methods perform well in many contexts [@lewis1998; @domingos1997; @rish2001]. This is likely because of the decision rule governing the final classifications---the posterior probabilities of the true class do not have to necessarily be statistically valid, they only need to be higher than those of every other class to be accurately classified.

Fortunately, as I will demonstrate, the predicted probabilities produced by my method are actually very well calibrated. This means that, unlike typical naive Bayes predicted probabilities, they accurately reflect the uncertainty in the algorithm's ethnorace predictions. Therefore researchers should feel confident in incorporating these measures of uncertainty into their analyses when using the method.

@@

To test the performance of the model, I apply the predictions to the combined North Carolina and Florida State voter file.[^12] These files contain snapshots of the registered voters in their respective states and provide individual-level data for first names, last names, address, political party, age, gender, and crucially self-identified ethnorace. After combining the two voter files, I then geocoded each unique address in the sample. This allowed me to match individual observations to Census places and blocks, and ZIP codes. Then I applied the prediction algorithm described above using the `bper` package and calculated each individuals' predicted ethnorace. In order to compare my method against an existing benchmark, I also used the `wru` package [@imai2016] to calculate ethnorace predictions for the same individuals.

[^12]: Retrieved May 2020

@@

## Calibration

As mentioned previously, the predicted probabilities produced by `bper` are remarkably well calibrated. This means that they accurately reflect the true probability that an individual belongs to a particular ethnorace. We can test this empirically by dividing the predicted probabilities for each ethnorace, for each individual in the sample, into evenly spaced bins (in this case 11). Then we compute the percentage of observed individuals of a particular ethnorace who fall in each bin. If the predicted probabilities are well calibrated, we would expect to see a one-to-one relationship between these two groups. For example, we should observe roughly 25% of individuals whose predicted probability of being White is in the neighborhood of 0.25 to actually be White.

Figure 1 displays the probability calibration plots for Asian, Black, Hispanic, and White predictions in the Florida/North Carolina sample. As demonstrated by their proximity to the dashed 45 degree line, the predictions for Black, Hispanic, and White individuals appear well calibrated. For each bin of predicted probabilities, we observe very close to the corresponding percentage of individuals of that ethnorace in the sample. Calibration for Asian predictions, on the other hand, appear slightly worse. Across the range of predicted probabilities, we observe between 5 to 20% fewer Asian individuals in each bin. This means that the algorithm has a tendency to overestimate Asian predicted probabilities. The comparatively poorer calibration for Asian Americans foreshadows some similar issues with predictive performance for that group in the next section.

```{r, fig.cap="Probability Calibration Plots by Ethnorace"}
cal_api <-
  ggplot(cal_data$bper_cal_api) +
  geom_line(color = "gray20") +
  geom_point(color = "gray20") +
  theme_minimal() +
  labs(y = "Observed Asian Percentage",
       x = "Predicted Probability Asian",
       title = "Asian") +
  ylim(0, 100) +
  theme(text = element_text(family = "serif"), plot.title = element_text(hjust = 0.5))

cal_black <-
  ggplot(cal_data$bper_cal_black) +
  geom_line(color = "gray20") +
  geom_point(color = "gray20") +
  theme_minimal() +
  labs(y = "Observed Black Percentage",
       x = "Predicted Probability Black",
       title = "Black") +
  ylim(0, 100) +
  theme(text = element_text(family = "serif"), plot.title = element_text(hjust = 0.5))

cal_hispanic <-
  ggplot(cal_data$bper_cal_hispanic) +
  geom_line(color = "gray20") +
  geom_point(color = "gray20") +
  theme_minimal() +
  labs(y = "Observed Hispanic Percentage",
       x = "Predicted Probability Hispanic",
       title = "Hispanic") +
  ylim(0, 100) +
  theme(text = element_text(family = "serif"), plot.title = element_text(hjust = 0.5))

cal_white <-
  ggplot(cal_data$bper_cal_white) +
  geom_line(color = "gray20") +
  geom_point(color = "gray20") +
  theme_minimal() +
  labs(y = "Observed White Percentage",
       x = "Predicted Probability White",
       title = "White") +
  ylim(0, 100) +
  theme(text = element_text(family = "serif"), plot.title = element_text(hjust = 0.5))

Rather than eye-balling plots like those in Figure 1, we can obtain objective measures of calibration by calculating Brier scores for each ethnorace. A Brier score is defined by$$BS = \frac{1}{N}\sum_{t=1}^N(f_t-o_t)^2$$Where $f_t$ is the predicted probability an individual $t$ belongs to a particular ethnorace and $o_t$ is the binary observed ethnorace value (1 if the individual belongs to a particular ethnorace, 0 otherwise). Brier scores range from 0 (good calibration) to 1 (bad calibration) and are analogous to a mean squared error. The Brier scores for the Florida/North Carolina sample are: Asian `r round(cal_data$bper_bs_api, 3)`; Black `r round(cal_data$bper_bs_black, 3)`; Hispanic `r round(cal_data$bper_bs_hispanic, 3)`; White `r round(cal_data$bper_bs_white, 3)`. These are all very low---indicating that the average predicted probability accurately reflects the uncertainty in ethnorace classification. The fact that the Brier score for Asian is lower than every other group might be puzzling given the calibration plots in Figure 1. However, this low score mostly reflects the abundance of low Asian predicted probabilities for individuals who are not Asian. The rarity of Asians in the Florida/North Carolina sample, therefore, produces a low Brier score for this group.

## Predictive Performance

While investigating the calibration of the predictions is important for assessing the reliability of the predicted probabilities, calibration does not tell us about how well the algorithm ranks these probabilities when generating final ethnorace predictions. Instead, the most straightforward metric for assessing individual-level predictive performance is the Accuracy score, or Overall Error Rate. This number is the proportion of correctly classified individuals in the sample. I ran models separately for different combinations of input variables to mimic data availability constraints in real-world applications, and then calculated the Accuracy score for each. Figure 2 displays a summary of the results of these different models.

@@

ggplot(acc_bper, aes(x = test_type, y = accuracy, group = 1)) +
  geom_line(color = "gray20") +
  geom_point(size = 3, color = "gray20") +
  theme_minimal() +
  ylim(0.75, 0.85) +
  labs(y = "Accuracy Score") +
  scale_x_discrete(name = "Model Inputs", limits = rev(c("all", "all_np", "zip", "zip_np", "place", "place_np", "county", "county_np", "state", "state_np")),
                   labels = rev(c(
                     "all" = "Block \n Multi-Unit \n Party ID \n First Name \n Last Name", 
                     "place" = "Place \n Party ID \n First Name \n Last Name", 
                     "county" = "County \n Party ID \n First Name \n Last Name", 
                     "all_np" = "Block \n Multi-Unit \n First Name \n Last Name", 
                     "place_np" = "Place \n First Name \n Last Name", 
                     "county_np" = "County \n First Name \n Last Name",
                     "zip" = "ZIP \n Party ID \n First Name \n Last Name",
                     "zip_np" = "ZIP \n First Name \n Last Name",
                     "state" = "State \n Party ID \n First Name \n Last Name",
                     "state_np" = "State \n First Name \n Last Name"
                   ))) +
  theme(text = element_text(family = "serif"), axis.text.x = element_text(size = 7))
```

As expected, the model with the greatest number of input data sources, and at the most precise geographic level, on the far right of the figure performs the best in terms of overall Accuracy. Using Census blocks, multi-unit address, party ID, first names, and last names, this model correctly identifies the ethnorace of `r round(subset(tests, select = "accuracy", test_method == "bper" & test_type == "all"), 3) * 100`% of individuals in the sample. The upward trend in model accuracy seen in the figure corresponds to shrinking the size of the geolocation variable used. Moving from state to county, from county to place, from place to ZIP code, and from ZIP code to block all improve the overall predictive performance in terms of overall Accuracy. I include the input for multi-unit occupancy only for the Census block models because I believe this reflects the practical contexts where `bper` might be used. If a researcher has access to individual-level addresses, they should be able to geocode these to find the matching Census blocks and should also be able to parse the residency type (multi-unit or stand-alone). But researchers relying on more aggregate geographies likely do not have access to individual addresses, and hence the residency characteristics, for their sample.[^13] In Figure 2 I pair each geolocation variable with a model missing the party ID input. The gains to predictive performance across all geographies appears roughly uniform with the addition of party ID.

[^13]: In the event that data is available, adding multi-unit occupancy inputs to aggregate geographies, such as ZIP codes, places, counties, or states *greatly* enhances the predictive accuracy of the model.

@@

ggplot(metrics_bper, aes(x = test_type, y = value, group = Metric)) +
  geom_line(aes(color = Metric), size = 0.3) +
  geom_point(aes(color = Metric), size = 2) +
  theme_minimal() +
  labs(x = "", y = "Score") +
  scale_color_manual(values = c("gray20", "gray50")) +
  scale_x_discrete(name = "Model Inputs", limits = rev(c("all", "all_np", "zip", "zip_np", "place", "place_np", "county", "county_np", "state", "state_np")),
                   labels = rev(c(
                     "all" = "Block", 
                     "place" = "Place", 
                     "county" = "County ", 
                     "all_np" = "Block (np)", 
                     "place_np" = "Place (np)", 
                     "county_np" = "County (np)",
                     "zip" = "ZIP",
                     "zip_np" = "ZIP (np)",
                     "state" = "State",
                     "state_np" = "State (np)"
                   ))) +
  facet_wrap(~ Race) +
  theme(text = element_text(family = "serif"), axis.text.x = element_text(size = 7, angle = 60))
```

Figure 3 displays the Precision and Recall scores broken down by ethnorace for the same models used in Figure 2.[^14] For both Whites and Hispanics, Precision and Recall remain high across all models. Using the most input data available (the models on the far right of the figure), White Precision and Recall are `r round(subset(tests, select = "prec_white", test_method == "bper" & test_type == "all"), 3)` and `r round(subset(tests, select = "rec_white", test_method == "bper" & test_type == "all"), 3)`, respectively. Hispanic Precision and Recall in this model are `r round(subset(tests, select = "prec_hispanic", test_method == "bper" & test_type == "all"), 3)` and `r round(subset(tests, select = "rec_hispanic", test_method == "bper" & test_type == "all"), 3)`, respectively. Hispanic predictive performance across all models likely remains high due to the distinctiveness of Spanish surnames. These metrics are uniformly lower for Asians---likely reflecting the relative rarity of Asian individuals in the sample. When using Census blocks as the unit of geography, predictions for African Americans are quite strong (Precision: `r round(subset(tests, select = "prec_black", test_method == "bper" & test_type == "all"), 3)`, Recall: `r round(subset(tests, select = "rec_black", test_method == "bper" & test_type == "all"), 3)`). But predictive performance falls with broader geolocations. At the state level, without party ID, Black Precision is `r round(subset(tests, select = "prec_black", test_method == "bper" & test_type == "state_np"), 3)` and Recall is `r round(subset(tests, select = "rec_black", test_method == "bper" & test_type == "state_np"), 3)`. This difference in predictive performance between geographies could be explained by the legacy of Black segregation in at least the North Carolina/Florida sample.

[^14]: The model names are abbreviated by the level of geography, and all use first name and last name inputs. Models with *(np)* do not use the party ID inputs.

@@

The classification metrics detailed above provide some information about the predictive shortcomings of my method. These limitations notwithstanding, however, the ethnorace predictions made by `bper` are nearly uniformly better than those from `wru`. Figure 4 displays the difference in Accuracy between `bper` and `wru` using the same input data discussed previously.[^15] The baseline of zero in the figure represents the Accuracy scores from each model using `wru`, and the height of the bars display the change in Accuracy using `bper`. Across all model types, `bper` scores roughly between 1.7 and 4.3 percentage points higher on this metric. Regardless of input data, `bper` classifies a higher proportion of individuals correctly in the North Carolina/Florida voter file.

[^15]: The `wru` package does not provide ZIP code or state level predictions so those models are excluded from the comparison.
